---
title: "acs1yr_income"
author: "MaryJoWebster"
date: "September 12, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, messages=FALSE, warning=FALSE}
#library(censusapi)
library(tidycensus)
library(tidyverse)
library(readxl)
library(janitor)
library(scales)
library(ggthemes)
library(leaflet)
library(sf)
library(stringr)
library(rmarkdown)
library(knitr)
library(reshape2)
library(DT)







#the first time you use your API key, you need to enter it here 
#this will load it into your REnviron file
#census_api_key("put your key here", install=TRUE)

#then on subsequent uses, you can just use this code
census_api_key(Sys.getenv("CENSUS_API_KEY"))

#this will generate a list of the tables and variables
#this particular code is asking for the 2017 5-year American community survey data
#it's storing it in a new data frame and then I've set it to write out to csv

#acs_variable_list <- load_variables(2017, "acs5", cache = TRUE)
#write.csv(acs_variable_list, 'acs_variable_list_2017.csv', row.names = FALSE)

my_variables <- c(all = "B19013_001",
                            black = "B19013B_001",
                            indian = "B19013C_001",
                            asian = "B19013D_001",
                            other = "B19013F_001",
                            multi = "B19013G_001",
                            white_nh = "B19013H_001",
                            hispanic = "B19013I_001",
                  one_per_HH = "B19019_002",
                            four_per_HH = "B19019_005",
                            med_earnings = "B20002_001",
                            med_earn_male = "B20002_002",
                            med_earn_female = "B20002_003")

years <-  lst(2012, 2013, 2014, 2015, 2016, 2017, 2018)


state_income <-  map_dfr(
  years,
  ~ get_acs(
    geography = "state",
    variables = my_variables,
    year = .x,
    survey = "acs1"
    ),
  .id = "year"
  ) %>% clean_names() %>% mutate(level='states')


county_income <-  map_dfr(
  years,
  ~ get_acs(
    geography = "county",
    variables = my_variables,
    state = "MN",
    year = .x,
    survey = "acs1"
    ),
  .id = "year"
  ) %>% clean_names() %>% mutate(level='mn counties')


us_income <-  map_dfr(
  years,
  ~ get_acs(
    geography = "us",
    variables = my_variables,
    year = .x,
    survey = "acs1"
    ),
  .id = "year"
  ) %>% clean_names() %>% mutate(level='us')

metro_income <-  map_dfr(
  years,
  ~ get_acs(
    geography = "metropolitan statistical area/micropolitan statistical area",
    variables = my_variables,
    year = .x,
    survey = "acs1"
    ),
  .id = "year"
  ) %>% clean_names() %>% mutate(level='metros')

#combine files

income <- bind_rows(county_income, state_income, us_income, metro_income)










```


```{r}

#clearly we need to adjust for inflation
#let's import CPI data
#be sure the Excel file is closed before running this
cpi <- read_excel("cpi_2000_2018.xlsx", sheet="BLS Data Series", range="A12:P31") %>% clean_names() %>% mutate(yr=as.character(year))

#pull out the CPI value for most current year that we'll use in the calculation
#row 15 for the 2014 CPI and row 19 for the 2018 CPI
#both are in column number 14
#pluck is part of purr package in tidyverse

cpi2018 <- cpi %>% pluck(14,19)

income <-  left_join(income, cpi %>% select(yr, annual), by=c("year"="yr")) %>% rename(cpi=annual)


#create a new column with the inflation-adjusted 
income <-  income %>% mutate(adj_estimate = (cpi2018/cpi)*estimate,
                             adj_moe = (cpi2018/cpi)*moe)


income <-  income %>% mutate(low=adj_estimate-adj_moe, high=adj_estimate+adj_moe)

```

```{r}
states_hhincome <-  income %>% filter(level=='states', variable=='all') %>% select(year, name, adj_estimate)

states_hhincome <-  dcast(states_hhincome, name ~ year)

write.csv(states_hhincome, './output/states_hhincome.csv', row.names=FALSE)



states_hhincome_moe <-  income %>% filter(level=='states', variable=='all') %>% select(year, name, adj_moe)

states_hhincome_moe <-  dcast(states_hhincome_moe, name ~ year)

write.csv(states_hhincome_moe, './output/states_hhincome_moe.csv', row.names=FALSE)
```

```{r}
states_earnings <-  income %>% filter(level=='states', variable=='med_earnings') %>% select(year, name, adj_estimate, adj_moe)

states_earnings_est <-  dcast(states_earnings %>% select(-adj_moe), name  ~ year)

states_earnings_moe <-  dcast(states_earnings %>% select(-adj_estimate), name  ~ year)

states_earnings_final <-  inner_join(states_earnings_est, states_earnings_moe, by=c("name"="name"))

write.csv(states_earnings_final, './output/states_earnings.csv', row.names=FALSE)


income %>% filter(level=='us', variable=='all')
```




#Median household income - 1 person households, adjusted
```{r}
income %>% filter(variable=='one_per_HH', name=='Minnesota') %>% select(year, adj_estimate, adj_moe) %>% mutate(low=adj_estimate-adj_moe, high=adj_estimate+adj_moe) %>% select(-adj_moe)
```

#4 person households
```{r}
income %>% filter(variable=='four_per_HH', name=='Minnesota') %>% select(year, adj_estimate, adj_moe) %>% mutate(low=adj_estimate-adj_moe, high=adj_estimate+adj_moe) %>% select(-adj_moe)
```



#Median household income -- all households, adjusted
```{r}
income %>% filter(variable=='all', name=='Minnesota') %>% select(year, adj_estimate, adj_moe) %>% mutate(low=adj_estimate-adj_moe, high=adj_estimate+adj_moe) %>% select(-adj_moe)
```

#Median earnings - both genders
```{r}
income %>% filter(variable=='med_earnings', name=='Minnesota') %>% select(year, adj_estimate, adj_moe) %>% mutate(low=adj_estimate-adj_moe, high=adj_estimate+adj_moe) %>% select(-adj_moe)
```

#Median earnings women
```{r}
income %>% filter(variable=='med_earn_female', name=='Minnesota') %>% select(year, adj_estimate, adj_moe) %>% mutate(low=adj_estimate-adj_moe, high=adj_estimate+adj_moe) %>% select(-adj_moe)
```

#Median earnings men
```{r}
income %>% filter(variable=='med_earn_male', name=='Minnesota') %>% select(year, adj_estimate, adj_moe) %>% mutate(low=adj_estimate-adj_moe, high=adj_estimate+adj_moe) %>% select(-adj_moe)
```

#Gender earnings gap
```{r}
pivot_wider(income %>% filter(name=='Minnesota', variable=='med_earn_male' | variable=='med_earn_female') %>% select(variable, year, adj_estimate), names_from = variable, values_from = adj_estimate ) %>% mutate(gap = med_earn_male-med_earn_female)
```




#State level, groups across time
```{r, message=FALSE, warning=FAL}

state_groups <- dcast(income%>% filter(name=='Minnesota') %>% select(year, variable, adj_estimate), variable ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



state_groups <- state_groups %>%  select(variable, y2014, y2015, y2016, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(state_groups)


state_groups_moe <- dcast(income%>% filter(name=='Minnesota') %>% select(year, variable, adj_moe), variable ~ year) 

state_groups_join <-  inner_join(state_groups, state_groups_moe, by=c("variable"="variable"))

write.csv(state_groups_join, './output/stategroups.csv')
```

#Median HH income -- all households -- by county

```{r, warning=FALSE, message=FALSE}

my_counties <-  c("27003", "27019", "27037", "27053", "27123", "27139", "27163", "27137",
                  "0500000US27003", "0500000US27019", "0500000US27037", "0500000US27053", "0500000US27123", "0500000US27139", "0500000US27163", "0500000US27137")



all_counties <- dcast(income%>% filter(variable=='all', geoid %in% my_counties) %>% select(year, name, adj_estimate), name ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



all_counties <- all_counties %>%  select(name, y2014, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(all_counties)
```




#Black households, by county
```{r, message=FALSE, warning=FALSE}

black_counties <- dcast(income%>% filter(variable=='black',geoid %in% my_counties) %>% select(year, name, adj_estimate), name ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



black_counties <- black_counties %>%  select(name, y2014, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(black_counties)
```


#Hispanic households, by county

```{r, message=FALSE, warning=FALSE}
hispanic_counties <- dcast(income%>% filter(variable=='hispanic',geoid %in% my_counties) %>% select(year, name, adj_estimate), name ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



hispanic_counties <- hispanic_counties %>%  select(name, y2014, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(hispanic_counties)
```

#Asian households, by county
```{r, message=FALSE, warning=FALSE}
asian_counties <- dcast(income%>% filter(variable=='asian',geoid %in% my_counties) %>% select(year, name, adj_estimate), name ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



asian_counties <- asian_counties %>%  select(name, y2014, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(asian_counties)
```

#4-person households, by county
```{r, message=FALSE, warning=FALSE}
fourper_counties <- dcast(income%>% filter(variable=='four_per_HH',geoid %in% my_counties) %>% select(year, name, adj_estimate), name ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



fourper_counties <- fourper_counties %>%  select(name, y2014, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(fourper_counties)
```

#Median earnings- females
```{r, message=FALSE, warning=FALSE}
female_counties <- dcast(income%>% filter(variable=='med_earn_female',geoid %in% my_counties) %>% select(year, name, adj_estimate), name ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



female_counties <- female_counties %>%  select(name, y2014, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(female_counties)
```

#Median earnings - males
```{r, warning=FALSE, message=FALSE}
male_counties <- dcast(income%>% filter(variable=='med_earn_male',geoid %in% my_counties) %>% select(year, name, adj_estimate), name ~ year) %>% rename(y2018=`2018`, y2017=`2017`, y2016=`2016`, y2015=`2015`, y2014=`2014`)



male_counties <- male_counties %>%  select(name, y2014, y2017,  y2018 )%>% mutate(y2014r= round(y2014,0), y2017r = round(y2017, 0), y2018r=round(y2018,0),diff= round((y2018r-y2014r)/y2014r*100,1)) %>% select(-y2014, -y2017,-y2018) %>% arrange(desc(diff))

datatable(male_counties)
```


```{r}

```


```{r}
income_pivot <-  income %>% filter(year=='2017' | year=='2018') %>% select(year, name, variable, adj_estimate, adj_moe)

write.csv(income_pivot, './output/income.csv', row.names=FALSE)

income_mn <-  pivot_wider(income %>% filter(name=='Minnesota') %>% select(year, variable, adj_estimate), names_from = year, values_from=adj_estimate)

write.csv(income_mn, './output/income_mn.csv', row.names=FALSE)

write.csv(income %>% filter(name=='United States', variable=='med_earnings') %>% select(year, adj_estimate, adj_moe), './output/us_earnings.csv', row.names=FALSE)
```

